### 2021-03-30 treatise-pdflatex/3_2_generating_beamforming_matrices.tex
Change dimension: context version
double check context version

In Haviv and Regev, ``The restricted isometry property of subsampled Fourier matrices'' \cite {KlM17}, Theorem \m {4.5} states a big O bound for the failure probability of a DFT submatrix with \m {\d_s} RIP.
We argue that their argument can be refined to give precise constants, by Chernoff-Hoeffding bounds and their theorem \m {2.2} with (say) \m {C \leftarrow 1/3}.
Their Theorem \m {3.7} and Theorem \m {4.5} may be changed accordingly, so that the relation below gives a bound for failure probability for \m {\d_s /N_{H,t} s} RIP.
In conclusion, (respectively for \m {\M {W} _R})
%
\DispNum {N:R:sd':NH} {
N_R
\geq \frac {s} {\d_s^2} \RB {\log \frac{1}{\d_s}}^2 \RB {\log \frac{s}{\d_s}}^2 \log N_{H,t}
}
%
Drop \m {\log \SB {1 /\d_s}^2} and the denominator \m {\d_s} in \m {\log \SB {s /\d_s}^2} in \eqref {N:R:sd':NH}, to get the statement below.

\Result
{Theorem}
{
Suppose
%
\DispNum {N:R:sd'NH} {
N_R
\geq \frac {s} {\d_s^2} \RB {\log s}^2 \log N_H 
}
%
Then \m {\M {F}_R} (respectively \m {\M {W} _R}) has \m {\d_s} RIP for probability \m {p}, with
\DispNum {1:p:ex:d's} {
1 -p
\leq \RB {\frac {\d_s} {N_H s}} ^{1/3} 
}
}

@@

We argue as follows.
If all angles \m {\th_l, \f_l} are exactly the multiples of \m {2 \pi/N_H}, then
\DispNum {G:G:Ud:dU} {
\M {G}
=\M {U} \mathrm {diag} \SB {\V {d}'} \M {U} ^\dagger 
}
for some \m {S}-sparse \m {\V {d}'} and unitary \m {\M {U}}.
Clearly,
\DispNum {G:2:g2:g2} {
\VNm {\M {G}} _2
=\VNm {\mathrm {abs} \SB {\V {g}}} _2 
}
where \m {\mathrm {abs}} denotes entrywise absolute value.
Then if we notice that \m {\M {W} _B \M {W} _R \M {K} \M {U}} and \m {\M {U} ^\dagger \M {K}^\dagger \M {F} _R^\dagger \M {F} _B^\Adj} has \m {\d_s} RIP too, we see
\DispNum {W:2:WB:d's} {
&\VNm {\M {W} _B \M {W} _R \M {H} \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U} \D
   \mathrm {diag} \SB {\mathrm {abs} \SB {\V {d}'}} \D
   \M {U} ^\dagger \M {K} ^\dagger \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}} \D
   \R {\mathrm {abs} \SB {\V {d}'}}
   \M {U} ^\dagger \M {K} ^\dagger \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}}} _2 \D
   \VNm {\M {F} _B ^\dagger \M {F} _R ^\dagger \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}}} _2 \notag \\
%
\eqsim &\VNm {\M {G}} _2 \RB {1 +2\d_s} 
}
Thus, even we cannot make sure that \m {\M {P}} is \m {\d_s} RIP with respect to all possible \m {\V {g}}'s, at least \m {\M {P}} has RIP with respect to relevant values of \m {\V {g}} according to the form \eqref {H:H:l0:ql}.

>>>>

In Haviv and Regev, ``The restricted isometry property of subsampled Fourier matrices'' \cite {KlM17}, Theorem \m {4.5} states a big O bound for the failure probability of a DFT submatrix with \m {\d_s} RIP.
We argue that their argument can be refined to give precise constants, by Chernoff-Hoeffding bounds and their theorem \m {2.2} with (say) \m {C = 1/3}.
Their Theorem \m {3.7} and Theorem \m {4.5} may be changed accordingly, so that the relation below gives a bound for failure probability for \m {\d_s /N_{H,t} s} RIP.
In conclusion, (respectively for \m {\M {W} _R})
%
\DispNum {N:R:sd':NH} {
N_R
\geq \frac {s} {\d_s^2} \RB {\log \frac{1}{\d_s}}^2 \RB {\log \frac{s}{\d_s}}^2 \log N_{H,t}
}
%
Drop \m {\log \SB {1 /\d_s}^2} and the denominator \m {\d_s} in \m {\log \SB {s /\d_s}^2} in \eqref {N:R:sd':NH}, to get the statement below.

\Result
{Theorem}
{
Suppose
%
\DispNum {N:R:sd'NH} {
N_R
\geq \frac {s} {\d_s^2} \RB {\log s}^2 \log N_{H,t}
}
%
Then \m {\M {F}_R} (respectively \m {\M {W} _R}) has \m {\d_s} RIP for probability \m {p}, with
\DispNum {1:p:ex:d's} {
1 -p
\leq \RB {\frac {\d_s} {N_{H,t} s}} ^{1/3} 
}
}

@@

We argue as follows.
If angles \m {\th_l} (respectively \m {\f_l}) are exactly the multiples of \m {2 \pi / N_{H,t}} (respectively \m {2 \pi / N_{H,r}}), then
\DispNum {G:G:Ud:dU} {
\M {G}
=\M {U} \mathrm {diag} \SB {\V {d}'} \M {U} ^\dagger 
}
for some \m {S}-sparse \m {\V {d}'} and unitary \m {\M {U}}.
Clearly,
\DispNum {G:2:g2:g2} {
\VNm {\M {G}} _2
=\VNm {\mathrm {abs} \SB {\V {g}}} _2 
}
where \m {\mathrm {abs}} denotes entrywise absolute value.
Then if we notice that \m {\M {W} _B \M {W} _R \M {K} \M {U}} and \m {\M {U} ^\dagger \M {K}^\dagger \M {F} _R^\dagger \M {F} _B^\Adj} has \m {\d_s} RIP too, we see
\DispNum {W:2:WB:d's} {
&\VNm {\M {W} _B \M {W} _R \M {H} \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U} \D
   \mathrm {diag} \SB {\mathrm {abs} \SB {\V {d}'}} \D
   \M {U} ^\dagger \M {K} ^\dagger \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}} \D
   \R {\mathrm {abs} \SB {\V {d}'}}
   \M {U} ^\dagger \M {K} ^\dagger \M {F} _R \M {F} _B} _2 \notag \\
%
= &\VNm {
   \M {W} _B \M {W} _R \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}}} _2 \D
   \VNm {\M {F} _B ^\dagger \M {F} _R ^\dagger \M {K} \M {U}
   \R {\mathrm {abs} \SB {\V {d}'}}} _2 \notag \\
%
\eqsim &\VNm {\M {G}} _2 \RB {1 +2\d_s} 
}
Thus, even we cannot make sure that \m {\M {P}} is \m {\d_s} RIP with respect to all possible \m {\V {g}}'s, at least \m {\M {P}} has RIP with respect to relevant values of \m {\V {g}} according to the form \eqref {H:H:l0:ql}.

### 2021-03-30 treatise-pdflatex/4_simulation.tex
double check context version

We use dimensionless noise level \m {\s} which takes value starting with \m {2^{-4}}, and being multiplied by powers of \m {2}, for \m {6} values.
We take \m {N_Y} to be \m {4, 6, 8}, respectively.
Three series of plots are simulated.
The first series for \m {N_{H,t} = 2 N_B} and \m {N_{H,r} = 2 N_B}.
The second series for \m {N_{H,t} = 2 N_B} and \m {N_{H,r} = 3 N_B}.
The third series for \m {N_{H,t} = 3 N_B} and \m {N_{H,r} = 2 N_B}.
Unfortunately, this is very far from achieving the ideal values \eqref {N:B:4l:H2}, and this may be part of the reason the result is not as successful as expected.

@@

For DS itself, first, \m {\tilde {\chi} _{\mathrm {DS}}} certainly decreases as \m {\s} increases, but it does not decrease indefinitely.
Second, \m {\tilde {\chi} _{\mathrm {DS}}} decreases as \m {N_H} increases.
Third, when \m {N_{B,t} /N_{B,r}} is smaller, \m {\tilde {\chi} _{\mathrm {DS}}} is also smaller, because we have a higher sample rate.

@@

\m {\chi} is not shown in the figure, because the big O bound is much larger than \m {\tilde {\chi}}.
It is possible that the nonsparsity of \m {\V {g}} undermines the analysis in chapter 3, despite our attempts to account for that effect.
It is curious to see whether for very large \m {N_H}, whether \m {\tilde {\chi}} and \m {\tilde {\chi}} will be asymptotically close.

In figures 24 and 25, we reproduce two cases of success and failure of DS.
In both figures the true values of \m {\M {H}} entries are ordered by magnitude, and corresponding estimated values in the first and second stage are plotted against them.
When DS fails, the soft thresholding did not rule out smaller entries, thus a second application of DS on the wrong index set failed too.
When DS succeeded, the soft thresholding gave a reasonable guess of the nonzero components, thus a second application of DS on this index set refined the estimated values.

@@

Suppose a linear program has an \m {N} dimensional variable and \m {M} inequality constraints, then its complexity is \m {\mathcal {O} \SB {N^2 M}}, assuming the Dantzig simplex method is used (\cite {BoV04}, p.6).
For our case, that would be \m {\mathcal {O} \SB {4 N_H ^4 \D 8 N_H^2} = \mathcal {O} \SB {N_H^6}}.
But other algorithms exist, where, in general, complexity of convex programs is difficult to analyze, as it is not easily to determine how many steps were calculated.

@@

Every Newton step involves a inversion of \m {\mathcal {O} \SB {\M {P} ^\dagger \M {P}}}, having the same order of magnitude to the time needed for matrix multiplication.
Since the dimension of \m {\M {P} ^\dagger \M {P}} is \m {N_H^2}, the cost of inversion may be roughly considered to be \m {\mathcal {O} \SB {N_H^4}}, as we know from newer results on the complexity of matrix inversion.
That is, we suppose \m {C_{\mathrm {step}} = N_H^4}.

We take some liberty to assume that, at the start the iteration, the Moore–Penrose inverse was taken to be the initial value of \m {\V {g}_0}, namely
%
\DispNum {g:0:gL:Py} {
\V {g}_0
=\V {g}_{\mathrm {LS}}
= \RB {\M {P} ^\dagger \M {P}} ^{-1} \M {P} ^\dagger \V {y} 
}
%
By above, we may write the complexity of DS to be
%
\DispNum {C:S:Og:g1} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {\VNm {\V {g}_{\mathrm {LS}} -\V {g} ^\star} _1} N_H^4 
}
%
Also assume
%
\DispNum {g:g:g;:;g} {
\V {g} ^\star
\approx \V {g} 
}
%
Indeed, this is the main thesis of our investigation.
And by restricted isometry, \m {\M {P}} has unity-normed, almost orthogonal columns, so
%
\DispNum {P:P:IN:H2} {
\M {P} ^\dagger \M {P}
\approx I _{N_{H}^2} 
}
%
Thus, by \eqref {g:0:gL:Py}, \eqref {C:S:Og:g1}, and \eqref {g:g:g;:;g},
%
\DispNum {C:S:Og:z1} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {
   \VNm {\V {g} +\RB {\M {P} ^\dagger \M {P}} ^{-1} \M {P} ^\dagger \V {z}
   -\V {g}} _1} N_H^4
=\mathcal {O} \SB {\VNm {\M {P} ^\dagger \V {z}} _1} N_H^4 
}

We see that \m {\RB {\M {P} ^\dagger \V {z}} _{\SB {j}}} observes complex standard normal distribution.
If we agree that, for high probability,
%
\DispNum {O:j:sO:O1} {
\mathcal {O} \SB {\VNm {\M {P} ^\dagger \V {z}}}
= \s N_{H,t} N_{H,r} \mathcal {O} \RB {1} 
}
Then we simply have
%
\DispNum {C:S:ON:H6} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {N_{H,t} ^3 N_{H,r} ^3} 
}

@@

Tropp and Gilbert \cite {TrG07a} discussed the complexity of OMP in the companion paper \cite {TrG07b}, giving
%
\DispNum {C:P:ON:NY} {
C_{\mathrm {OMP}}
=\mathcal {O} \SB {N_{H,t} N_{H,r} \log N_B} 
}
%
Here \m {\log \RB {N_{B,r} N_{B,t}}} is small in our experiments, and if we dropped it, we may take
\DispNum {C:P:ON:H2} {
C_{\mathrm {OMP}}
=\mathcal {O} \SB {N_{H,t} N_{H,r}}
}

@@

Lasso is equivalent to a quadratically constrained quadratic program, and there is not always a closed form for its complexity.
However, Lasso and DS are equivalent in certain conditions \cite {AsR10}, and we may suppose here they have the same complexity, \m {\mathcal {O} \SB {N_{H,t} ^3 N_{H,r} ^3}}.

@@

Therefore we take
\DispNum {C:o:ON:H6} {
C_{\mathrm {Lasso}}
=\mathcal {O} \SB {N_{H,t} ^3 N_{H,r} ^3}
}

@@

Abbreviate \m {N_h = N_{H,t} \D N_{H,r}} as in the last chapter.
Use \m {t_{\mathrm {OMP}} \SB {N_h}}, \m {t_{\mathrm {Lasso}} \SB {N_h}}, \m {t_{\mathrm {DS}} \SB {N_h}} to denote their respective time taken.
We here copy the arithmetic average (including different \m {\s}) of simulated values below, in seconds.
They are:
\m {t_{\mathrm {OMP}} \SB {12}  = 0.0119,}
\m {t_{\mathrm {OMP}} \SB {16} = 0.0297, }
\m {t_{\mathrm {OMP}} \SB {20}  = 0.0600,}
\m {t_{\mathrm {OMP}} \SB {24} = 0.155, }
%
\m { t_{\mathrm {Lasso}} \SB {12}  = 0.0436,}
\m { t_{\mathrm {Lasso}} \SB {16} = 0.176, }
\m { t_{\mathrm {Lasso}} \SB {20}  = 0.471,}
\m { t_{\mathrm {Lasso}} \SB {24} = 1.26, }
%
\m { t_{\mathrm {DS}} \SB {12}  = 0.357,}
\m { t_{\mathrm {DS}} \SB {16} = 2.07, }
\m { t_{\mathrm {DS}} \SB {20}  = 29.8,}
\m { t_{\mathrm {DS}} \SB {24} = 91.5}

Simulation illustrates that \m {t_{\mathrm {DS}}} grows tremendously with \m {N_h}, agreeing Friedlander and Saunders \cite {FrS07} who noted the high complexity of DS.
This is not surprising, since the inversion of \m {\M {P} \M {P} ^\Adj} is probably the bottleneck of complexity, which has dimension \m {\M {N}_H^2}.
By the way, different tolerance levels and maximal iteration numbers may result in vastly different computation time.

@@

\DispNum {t:P:NH:83} {
\tilde {t}_{\mathrm {OMP}} \eqsim &\mathcal {O} \SB {N_h^{3.6}} \\
\tilde {t}_{\mathrm {Lasso}} \eqsim &\mathcal {O} \SB {N_h^{4.8}} \\
\tilde {t}_{\mathrm {DS}} \eqsim &\mathcal {O} \SB {N_h^{8.3}} 
}

>>>>

We use dimensionless noise level \m {\s} which takes value starting with \m {2^{-4}}, and being multiplied by powers of \m {2}, for \m {6} values.
We take \m {N_B} to be \m {4, 6, 8}, respectively.
Three series of plots are simulated.
The first series for \m {N_{H,t} = 2 N_B} and \m {N_{H,r} = 2 N_B}.
The second series for \m {N_{H,t} = 2 N_B} and \m {N_{H,r} = 3 N_B}.
The third series for \m {N_{H,t} = 3 N_B} and \m {N_{H,r} = 2 N_B}.
Unfortunately, this is very far from achieving the ideal values \eqref {N:B:4l:H2}, and this may be part of the reason the result is not as successful as expected.

@@

For DS itself, first, \m {\tilde {\chi} _{\mathrm {DS}}} certainly decreases as \m {\s} increases, but it does not decrease indefinitely.
Second, \m {\tilde {\chi} _{\mathrm {DS}}} decreases as \m {N_H} increases.

@@

\m {\chi} is not shown in the figure, because the big O bound is much larger than \m {\tilde {\chi}}.
It is possible that the nonsparsity of \m {\V {g}} undermines the analysis in chapter 3, despite our attempts to account for that effect.
It is curious to see whether for very large \m {N_H}, whether \m {\tilde {\chi}} and \m {\chi} will be asymptotically close.

In the below, we reproduce two cases of success and failure of DS.
In both figures the true values of \m {\M {H}} entries are ordered by magnitude, and corresponding estimated values in the first and second stage are plotted against them.
When DS fails, the soft thresholding did not rule out smaller entries, thus a second application of DS on the wrong index set failed too.
When DS succeeded, the soft thresholding gave a reasonable guess of the nonzero components, thus a second application of DS on this index set refined the estimated values.

@@

Suppose a linear program has an \m {N} dimensional variable and \m {M} inequality constraints, then its complexity is \m {\mathcal {O} \SB {N^2 M}}, assuming the Dantzig simplex method is used (\cite {BoV04}, p.6).
For our case, that would be \m {\mathcal {O} \SB {4 N_h ^2 \D 8 N_h} = \mathcal {O} \SB {N_h ^3}}.
But other algorithms exist, where, in general, complexity of convex programs is difficult to analyze, as it is not easily to determine how many steps were calculated.

@@

Every Newton step involves a inversion of \m {\mathcal {O} \SB {\M {P} ^\dagger \M {P}}}, having the same order of magnitude to the time needed for matrix multiplication.
Since the dimension of \m {\M {P} ^\dagger \M {P}} is \m {N_h \D N_h}, the cost of inversion may be roughly considered to be \m {\mathcal {O} \SB {N_h^2}}, as we know from newer results on the complexity of matrix inversion.
That is, we suppose \m {C_{\mathrm {step}} = N_h^2}.

We take some liberty to assume that, at the start the iteration, the Moore–Penrose inverse was taken to be the initial value of \m {\V {g}_0}, namely
%
\DispNum {g:0:gL:Py} {
\V {g}_0
=\V {g}_{\mathrm {LS}}
= \RB {\M {P} ^\dagger \M {P}} ^{-1} \M {P} ^\dagger \V {y} 
}
%
By above, we may write the complexity of DS to be
%
\DispNum {C:S:Og:g1} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {\VNm {\V {g}_{\mathrm {LS}} -\V {g} ^\star} _1} N_h^2 
}
%
Also assume
%
\DispNum {g:g:g;:;g} {
\V {g} ^\star
\approx \V {g} 
}
%
Indeed, this is the main thesis of our investigation.
And by restricted isometry, \m {\M {P}} has unity-normed, almost orthogonal columns, so
%
\DispNum {P:P:IN:H2} {
\M {P} ^\dagger \M {P}
\approx I _{N_h} 
}
%
Thus, by \eqref {g:0:gL:Py}, \eqref {C:S:Og:g1}, and \eqref {g:g:g;:;g},
%
\DispNum {C:S:Og:z1} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {
   \VNm {\V {g} +\RB {\M {P} ^\dagger \M {P}} ^{-1} \M {P} ^\dagger \V {z}
   -\V {g}} _1} N_h ^2
=\mathcal {O} \SB {\VNm {\M {P} ^\dagger \V {z}} _1} N_h ^2 
}

We see that \m {\RB {\M {P} ^\dagger \V {z}} _{\SB {j}}} follows complex standard normal distribution.
If we agree that, for high probability,
%
\DispNum {O:j:sO:O1} {
\mathcal {O} \SB {\VNm {\M {P} ^\dagger \V {z}}}
= \s N_h \mathcal {O} \RB {1} 
}
Then we simply have
%
\DispNum {C:S:ON:H6} {
C_{\mathrm {DS}}
=\mathcal {O} \SB {N_h ^3} 
}

@@

Tropp and Gilbert \cite {TrG07a} discussed the complexity of OMP in the companion paper \cite {TrG07b}, giving
%
\DispNum {C:P:ON:NY} {
C_{\mathrm {OMP}}
=\mathcal {O} \SB {L \RB {\log N_h} ^2} 
}

@@

Lasso is equivalent to a quadratically constrained quadratic program, and there is not always a closed form for its complexity.
However, Lasso and DS are equivalent in certain conditions \cite {AsR10}, and we may suppose here they have the same complexity, \m {\mathcal {O} \SB {N_h ^3}}.

@@

Therefore we take
\DispNum {C:o:ON:H6} {
C_{\mathrm {Lasso}}
=\mathcal {O} \SB {N_h ^3}
}

@@

Simulation illustrates that \m {t_{\mathrm {DS}}} grows tremendously with \m {N_h}, agreeing Friedlander and Saunders \cite {FrS07} who noted the high complexity of DS.
This is not surprising, since the inversion of \m {\M {P} \M {P} ^\Adj} is probably the bottleneck of complexity, which has dimension \m {N_h}.
By the way, different tolerance levels and maximal iteration numbers may result in vastly different computation time.

@@

\DispNum {t:P:NH:83} {
\tilde {t}_{\mathrm {OMP}} \eqsim &\mathcal {O} \SB {N_h^{1.8}} \\
\tilde {t}_{\mathrm {Lasso}} \eqsim &\mathcal {O} \SB {N_h^{2.4}} \\
\tilde {t}_{\mathrm {DS}} \eqsim &\mathcal {O} \SB {N_h^{4.2}} 
}
